## Create database 

```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Below are the instructions and code for getting data from wildtrax that was developed by Elly Knight. The original rScripts are available [here](https://github.com/baynelab-research/aru-data-processing-code).

*See the wildRtrax article here for more details on authentication:* <https://abbiodiversity.github.io/wildRtrax/articles/authenticating-into-wt.html>

1.  **Load packages**

```{r eval=FALSE}
#install.packages("remotes")
#remotes::install_github("ABbiodiversity/wildRtrax")
library(wildRtrax)
library(tidyverse)
library(lubridate)
library(unmarked)
```


2.  **Login to WildTrax**

NOTE: If you have not done this yet, first, open a blank R document and add the following text to it:  
```{r eval=FALSE}
 Sys.setenv(WT_USERNAME = 'your wildtrax username', WT_PASSWORD = 'your wildtrax password')  
```

Then save the file to "1_scripts/functions/login.R", and ensure that it will not be pushed to github by adding the file path (without the quotes) to your .gitignore file (in the main project directory)

```{r eval=FALSE}
config <- "1_scripts/functions/login.R"
source(config)
wt_auth()
```

3.  **Get list of projects from WildTrax**

```{r eval=FALSE}
projects <- wt_get_download_summary(sensor_id = 'ARU')
```

4.  **Download RUGR dataset summary report** 
This takes a little while 

```{r eval=FALSE}
# Find the project_id for the RUGR project
rugr_id <- projects[grep("RUGR", projects$project), "project_id"][[1]]
# project_id = 1321

dat.rugr <- wt_download_report(project_id = rugr_id, sensor_id = 'ARU', weather_cols = T, report = "summary")

saveRDS(dat.rugr, paste0("0_data/raw/dat-rugr_", format(Sys.time(), "%d-%B-%Y"), ".rds")) 
```

5.  **Download RUGR task report to check coordinate buffering**

```{r eval=FALSE}
task.rugr <- wt_download_report(project_id = rugr_id, sensor_id = 'ARU', report = "task")
table(task.rugr$buffer)
```


6.  **Format the data for occupancy estimation in unmarked**  
```{r, eval = FALSE}

# Load the data if not already in the environment
dat.rugr <- readRDS("0_data/raw/dat-rugr_25-February-2023.rds")

# Call the function to format the data
source("1_scripts/functions/wt_format_occupancy.R") 

# Filter the data to only the surveys that have been transcribed 
dat <- dat.rugr %>% filter(status == "Transcribed")

# Run the data through the function 
occ.data <- wt_format_occupancy(data = dat.rugr, species = 'RUGR')

# Check the distributions of day of year and time of day
hist(occ.data@obsCovs$doy)
hist(occ.data@obsCovs$hr)
``` 
Because the data come from 2 different time-of-day clusters (nocturnal and diurnal), the detection model will not perform well if we use both clusters in a standard linear modelling format. Therefore, we will use only the diurnal surveys. 
```{r, eval=FALSE}
# Reformat the data to only include crepuscular and diurnal surveys (those done after 02:00)
dat1 <- dat %>%
    dplyr::filter(species_code=='RUGR') %>%
    dplyr::select(location, recording_date) %>%
    mutate(occur=1) %>%
    right_join(dat %>%
                 dplyr::select(location, recording_date, observer, method, species_code) %>%
                 unique(),
               by=c("location", "recording_date")) %>%
    mutate(occur = ifelse(is.na(occur), 0, 1),
           doy = yday(recording_date),
           hr = as.numeric(hour(recording_date) + minute(recording_date)/60))

dat.day <- dat1 %>% filter(hr > 2)

occ.data <- wt_format_occupancy(data = dat.day, species = 'RUGR')

# Divide day or year by 100 to make the ML estimation easier (and more reliable)
occ.data@obsCovs$doy <- occ.data@obsCovs$doy/100
occ.data@obsCovs$doy2 <- (occ.data@obsCovs$doy)^2

# Or, alternatively, standardize all of the detection covariates 
occ.data.1 <- occ.data
occ.data.1@obsCovs$doy <- scale(occ.data@obsCovs$doy)
occ.data.1@obsCovs$doy2 <- scale(occ.data@obsCovs$doy2)
occ.data.1@obsCovs$hr <- scale(occ.data.1@obsCovs$hr)
occ.data.1@obsCovs$hr2 <- scale(occ.data.1@obsCovs$hr2)

# Run the basic occupancy model with covariates on detection but assuming constant occupancy
occ1 <- occu(~ doy + doy2 + hr ~ 1, occ.data) 

occ1.1 <- occu(~ doy + doy2 + hr + hr2 ~ 1, occ.data.1)

backTransform(occ1, type = "state")
``` 

Plot the effects of day of year and time of day on detection probability to see if it makes sense 
```{r, eval=FALSE}
det.doy <- data.frame(doy = seq(1.1, 1.5, 0.01), doy2 = seq(1.1, 1.5, 0.01)^2, hr = mean(occ.data@obsCovs$hr, na.rm = TRUE), hr2 = mean(occ.data@obsCovs$hr2, na.rm = TRUE))
pred.doy <- predict(occ1, newdata = det.doy, type = "det") 

plot(det.doy$doy, pred.doy$Predicted)

det.hr <- data.frame(doy = mean(occ.data@obsCovs$doy, na.rm = TRUE), doy2 = mean(occ.data@obsCovs$doy2, na.rm = TRUE), hr = seq(5, 8, 0.1))

pred.hr <- predict(occ1, newdata = det.hr, type = "det") 

plot(det.hr$hr, pred.hr$Predicted)

```





