## Create database 

```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Below are the instructions and code for getting data from wildtrax that was developed by Elly Knight. The original rScripts are available [here](https://github.com/baynelab-research/aru-data-processing-code).

*See the wildRtrax article here for more details on authentication:* <https://abbiodiversity.github.io/wildRtrax/articles/authenticating-into-wt.html>

1.  **Load packages**

```{r eval=FALSE}
#install.packages("remotes")
#remotes::install_github("ABbiodiversity/wildRtrax")
library(wildRtrax)
library(tidyverse)
```


2.  **Login to WildTrax**

NOTE: If you have not done this yet, first, open a blank R document and add the following text to it:  
```{r eval=FALSE}
 Sys.setenv(WT_USERNAME = 'your wildtrax username', WT_PASSWORD = 'your wildtrax password')  
```

Then save the file to "1_scripts/functions/login.R", and ensure that it will not be pushed to github by adding the file path (without the quotes) to your .gitignore file (in the main project directory)

```{r eval=FALSE}
config <- "login.R"
source(config)
wt_auth()
```

3.  **Get list of projects from WildTrax**

```{r eval=FALSE}
projects <- wt_get_download_summary(sensor_id = 'ARU')
```

4.  **Download RUGR dataset summary report**

```{r eval=FALSE}
dat.rugr <- wt_download_report(project_id = 1321, sensor_id = 'ARU', weather_cols = T, report = "summary")
```

5.  **Download RUGR task report to check coordinate buffering**

```{r eval=FALSE}
task.rugr <- wt_download_report(project_id = 1321, sensor_id = 'ARU', report = "task")
table(task.rugr$buffer)
```

